Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-03-29 cis-perf-new-ll.tex
More on new version

For a detected episode $j$, we observe $O_j = [l_j^{(b)}, r_j^{(b)}, l_j^{(e)}, r_j^{e}, i(j)]^T$.
These are: the individual in which the episode occurs, denoted $i(j)$; the earliest and latest time the episode could have begun, $l_j^{(b)}$ and $r_j^{(b)}$ respectively; and the earliest and latest time the episode could have ended, $l_j^{(e)}$ and $r_j^{(e)}$ respectively.

For an undetected episode $j$, there are no observations.
In which case, define $O(j) = \emptyset$.

Define an integer $D$ and $\set{D} = \vec{\nu}_1, \dots, \vec{\nu}_D$ as the set of all possible observations of detected episodes; that is, $O(j) \in \set{D}$ if and only if $j$ is a detected infection.
For any $\vec{\nu} = [l^{(b)}, r^{(b)}, l^{(e)}, r^{e}, i]^T$, $\vec{\nu} \in \set{D}$ if and only if the following conditions hold.
\begin{enumerate}
  \item $l^{(b)} - 1, r^{(b)}, l^{(e)}, r^{e} + 1 \in \sched_i$, that is individual $i$ was tested at all these times.
  \item $r^{(b)} = \min\{t \in \sched_i \ssep t \geq l^{(b)}\}$, that is $r^{(b)}$ is the first time at or after $l^{(b)}$ that individual $i$ was tested.
  \item $r^{(e)} = \min\{t \in \sched_i \ssep t > l^{(e)}\} - 1$, that is $r^{(e)}$ is the day before the next time after $l^{(e)}$ that individual $i$ was tested.
  \item $1 \leq r^{(b)} \leq T$, that is the episode begins in the period of interest.
\end{enumerate}
Let $n_k$ denote the number of observations of $\vec{\nu}_k$, $\nnodet$ the latent number of undetected episodes, $\vec{n} = [n_1, \dots, n_D, \nnodet]^T$, and $\ntot = \sum_{i=1}^D n_i + \nnodet$.
For $\vec{\nu}_k \in \set{D}$, let $p_k = \prob(O(j) = \vec{\nu}_k \mid \vec{\theta})$, $p_u = \prob(O(j) = \emptyset \mid \vec{\theta})$, and $\vec{p} = [p_1, \dots, p_D, p_u]^T$.
Assume, for tractability, that the events $O(j) = \vec{\nu}_k$ and $O(j') = \vec{\nu}_{k'}$ are independent for $j \neq j'$; this assumption is discussed in \cref{perf-test:sec:discussion}.
Then:
\begin{align}
  \vec{n} \mid \ntot, \vec{\theta} &\dist \MN(\ntot, \vec{p})
\intertext{that is:}
  p(\vec{n} \mid \ntot, \vec{\theta}) &= \frac{\ntot!}{\nnodet!\prod_{k=1}^D n_k!} p_u^{\nnodet} \prod_{k=1}^D p_k^{n_k}.
  \label{perf-test:eq:multinomial-ll}
\end{align}

In the CIS data, each $n_k$ is either 0 or 1.
Let $k \in \set{C}$ indicate that $n_k = 1$.
Furthermore, note that the support of the multinomial distribution requires that $\nnodet = \ntot - \ndet$.
Then \cref{perf-test:eq:multinomial-ll} simplifies to:
\begin{align}
  p(\vec{n} \mid \ntot, \vec{\theta})
  &= \frac{\ntot!}{(\ntot - \ndet)!} p_u^{\ntot-\ndet} \prod_{k \in \set{C}} p_k.
  \label{perf-test:eq:multinomial}
\end{align}

@@

>>>>

For any episode $j$, define $O(j)$ as the observations of that episode.
Test results are deterministic based on the time of the test and the individual's infection status.
Therefore, $O(j)$ is fully determined by the triplet $(B_j, E_j, i(j))$, $j$'s beginning time, end time, and the individual in which the episode occurs respectively.

For a detected episode, we have $O(j) = [l_j^{(b)}, r_j^{(b)}, l_j^{(e)}, r_j^{e}, i(j)]^T$ (see \cref{E-biology-data:sec:cis-episodes}).
These are: the individual in which the episode occurs, denoted $i(j)$; the earliest and latest time the episode could have begun, $l_j^{(b)}$ and $r_j^{(b)}$ respectively; and the earliest and latest time the episode could have ended, $l_j^{(e)}$ and $r_j^{(e)}$ respectively.

For an undetected episode $j$, there are no observations; in which case, define $O(j) = \emptyset$.

Define an integer $N_D$ and $\set{D} = \vec{\nu}_1, \dots, \vec{\nu}_{N_D}$ as the set of all possible observations of detected episodes; that is, $O(j) \in \set{D}$ if and only if $j$ is a detected infection.
For any $\vec{\nu} = [l^{(b)}, r^{(b)}, l^{(e)}, r^{e}, i]^T$, $\vec{\nu} \in \set{D}$ if and only if all the following conditions hold.
\begin{enumerate}
  \item $l^{(b)} - 1, r^{(b)}, l^{(e)}, r^{e} + 1 \in \sched_i$, that is individual $i$ was tested at all these times.
  \item $l^{(b)} \leq r^{(b)} \leq l^{(e)} \leq r^{(e)}$.
  \item $r^{(b)} = \min\{t \in \sched_i \ssep t \geq l^{(b)}\}$, that is $r^{(b)}$ is the first time at or after $l^{(b)}$ that individual $i$ was tested.
  \item $r^{(e)} = \min\{t \in \sched_i \ssep t > l^{(e)}\} - 1$, that is $r^{(e)}$ is the day before the next time after $l^{(e)}$ that individual $i$ was tested.
  \item $1 \leq r^{(b)} \leq T$, that is the episode begins in the period of interest.
\end{enumerate}
Let $n_k$ denote the number of observations of $\vec{\nu}_k$, $\nnodet$ the latent number of undetected episodes, $\vec{n} = [n_1, \dots, n_D, \nnodet]^T$, and $\ntot = \sum_{i=1}^{N_D} n_i + \nnodet$.
For $\vec{\nu}_k \in \set{D}$, let $p_k = \prob(O(j) = \vec{\nu}_k \mid \vec{\theta})$, $p_u = \prob(O(j) = \emptyset \mid \vec{\theta})$, and $\vec{p} = [p_1, \dots, p_D, p_u]^T$.
Assume, for tractability, that the events $O(j) = \vec{\nu}_k$ and $O(j') = \vec{\nu}_{k'}$ are independent for $j \neq j'$; this assumption is discussed in \cref{perf-test:sec:discussion}.
Then:
\begin{align}
  \vec{n} \mid \ntot, \vec{\theta} &\dist \MN(\ntot, \vec{p})
\intertext{that is:}
  p(\vec{n} \mid \ntot, \vec{\theta}) &= \frac{\ntot!}{\nnodet!\prod_{k=1}^{N_D} n_k!} p_u^{\nnodet} \prod_{k=1}^{N_D} p_k^{n_k}.
  \label{perf-test:eq:multinomial-ll}
\end{align}

In the CIS data, each $n_k$ ($k \neq u$) is observed as either 0 or 1.
Let $k \in \set{C}$ indicate that $n_k = 1$.
Furthermore, note that the support of the multinomial distribution requires that $\nnodet = \ntot - \ndet$.
Then \cref{perf-test:eq:multinomial-ll} simplifies to:
\begin{align}
  p(\vec{n} \mid \ntot, \vec{\theta})
  &= \frac{\ntot!}{(\ntot - \ndet)!} p_u^{\ntot-\ndet} \prod_{k \in \set{C}} p_k.
  \label{perf-test:eq:multinomial}
\end{align}

@@

Introducing false negatives means that $O(j)$ is now a random variable because the test results between the start and the end of the episode could contain false negatives.
Therefore, the
